\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{epigraph} 

\title{Linear Algebra Done Right\\Solutions to Exercises 6.A}
\author{}
\date{}

\providecommand{\abs}[1]{\lvert#1\rvert} \providecommand{\norm}[1]{\lVert#1\rVert}

\begin{document}

\maketitle

\section{Inhomogeneous inner products}
\subsection*{Problem statement}
Show that the function that takes $\big((x_1,x_2),(y_1,y_2)\big)\in \mathbf{R}^2\times\mathbf{R}^2$ to $|x_1y_1|+|x_2y_2|$ is not an inner product on $\mathbf{R}^2$.

\subsection*{Solution}
The inner product is not homogeneous (Definition 6.3). 
Consider $\lambda=-2$. 
\begin{align*} 
\langle(-2x_1,-2x_2),(y_1,y_2)\rangle &= |-2x_1y_1|+|-2x_2y_2| \\ 
 &= 2|x_1y_1| + 2|x_2y_2|\\
 &= 2\langle(x_1,x_2),(y_1,y_2)\rangle\\
 &\neq -2\langle(x_1,x_2),(y_1,y_2)\rangle
\end{align*}

\clearpage

\section{Indefinite inner products}
\subsection*{Problem statement}
Show that the function that takes $\big((x_1,x_2,x_3),(y_1,y_2,y_3)\big)\in \mathbf{R}^3\times\mathbf{R}^3$\newline to $x_1y_1+x_3y_3$ is not an inner product on $\mathbf{R}^3$.

\subsection*{Solution}
The inner product does not obey definiteness (Definition 6.3). 
Consider\newline $(x_1,x_2,x_3)=(y_1,y_2,y_3)=(0,1,0)$.
\[\langle(x_1,x_2,x_3),(y_1,y_2,y_3)\rangle=\langle(0,1,0),(0,1,0)\rangle=0\cdot 0+0\cdot 0=0\]

\clearpage

\section{Replacing the positivity condition}
\subsection*{Problem statement}
Suppose $\mathbf{F}=\mathbf{R}$ and $V\neq \{0\}$. 
Replace the positivity condition (which states that $\langle v, v\rangle\geq 0$ and for all $v\in V$) in the definition of an inner product (6.3) with the condition that $\langle v, v\rangle > 0$ for some $v\in V$. 
Show that this change in the definition does not change the set of functions from $V\times V$ to $\mathbf{R}$ that are inner products on $V$.

\subsection*{Solution}
The set of functions from $V\times V$ to $\mathbf{R}$ that are inner products on $V$ clearly does not grow smaller. 
Thus we must show that the set does not grow larger.

Suppose $v\in V$ is such that it satisfies the ``new'' condition that $\langle v, v\rangle > 0$. 
There are two cases to consider: $V=\operatorname{span}(v)$ and $V\neq\operatorname{span}(v)$.
\\
\\
$\mathbf{{V=\operatorname{\textbf{span}}(v)}}$: For any vector $u\in V$ there exists $\lambda\in\mathbf{R}$ such that $u=\lambda v$. 
Hence 
\[\langle u, u\rangle =\langle \lambda v, \lambda v \rangle = \abs{\lambda}^2\langle v, v\rangle \]
Given $\langle v, v\rangle > 0$ and $|\lambda|^2\geq 0$ for all $\lambda\in\mathbf{R}$, it follows that 
\[\langle u, u\rangle = |\lambda|^2\langle v, v\rangle \geq 0\]
Thus we've shown that the positivity condition is fulfilled and the set of function that are inner products on $V$ clearly does not grow larger.
\\
\\
$\mathbf{V\neq\operatorname{\textbf{span}}(v)}$: Since $V\neq \{0\}$ and $V\neq\operatorname{span}(v)$, there exists $u\in V$ that is not a scalar multiple of $v$. 
Theorem 6.14 (`An orthogonal decomposition') allows us to decompose $u$ into $u=cv+w$ where $v$ and $w$ are orthogonal. 
From the Pythagorean theorem, it follows
\[\norm{v+w}^2=\norm{v}^2+\norm{w}^2\]

Let's assume that $||w||^2 < 0$. 
Now consider the vector $\lambda w$ where $\lambda=\frac{\norm{v}}{\norm{w}}$\footnote{
The $\abs{\frac{\norm{v}}{\norm{w}}}^2 \norm{w}^2$ term is quite harry, but note that $\norm{w}^2 < 0$ and $\abs{\norm{w}}^2 > 0$, therefore $\frac{\norm{w}^2}{\abs{\norm{w}}^2} = -1 $. Also note $\abs{\norm{v}}^2 = \norm{v}^2$ since $\langle v, v\rangle > 0$.
}
\[\norm{v+\lambda w}^2=\norm{v}^2+\norm{\lambda w}^2 = \norm{v}^2+\abs{\frac{\norm{v}}{\norm{w}}}^2 \norm{w}^2 = \norm{v}^2 - \norm{v}^2 = 0 \]
Thus we have a contradiction since $v+\lambda w\neq0$ yet $||v+\lambda w||^2=0$, defying the property of definiteness (Definition 6.3). Therefore $||w||^2 \geq 0$.

It then follows that 
\[||u||^2 = ||v+w||^2=||v||^2+||w||^2 \geq 0\]
showing that the positivity condition is fulfilled and the set of function that are inner products on $V$ does not grow larger.

\clearpage

\section{Diagonals of a rhombus are perpendicular}
\subsection*{Problem statement}
Suppose $V$ is a real inner product space.
\begin{itemize}
    \item[(a)] Show that $\langle u+v,u-v\rangle=\norm{u}^2 - \norm{v}^2$ for every $u,v\in V$.
    \item[(b)] Show that if $u,v\in V$ have the same norm, then $u+v$ is orthogonal to $u-v$.
    \item[(c)] Use part (b) to show that the diagonals of a rhombus are perpendicular to each other.
\end{itemize}

\subsection*{Solution}

\subsubsection*{a}
Via the definition of the inner product (Definition 6.3) and Theorem 6.7 (`Basic properties of an inner product'), we can expand $\langle u+v,u-v\rangle$ to
\[\langle u+v,u-v\rangle=\langle u,u\rangle+\langle v,u\rangle + \langle u,-v\rangle + \langle v,-v\rangle.\]

Since $V$ is a real inner product space, we have
\[\langle u,-v\rangle=-1\langle u,v\rangle \;\;\;\text{and}\;\;\; \langle v,-v\rangle=-\norm{v}^2\]
and since $\langle u,v\rangle=\overline{\langle v,u\rangle}=\langle v,u \rangle$ for reals, we can write
\[\langle u+v,u-v\rangle=\norm{u}^2+\langle v,u\rangle -\langle v,u \rangle - \norm{v}^2=\norm{u}^2 - \norm{v}^2.\]

\subsubsection*{b}
If $\norm{u} = \norm{v}$, then $\norm{u}^2 = \norm{v}^2$. 
Hence, via part (a), we can write
\[\langle u+v,u-v\rangle=\norm{u}^2 - \norm{v}^2=\norm{u}^2 -\norm{u}^2=0\]
and $u+v$ is orthogonal to $u-v$.

\subsubsection*{c}
In a rhombus, all sides are the same length, and hence, have the same norm. 
Suppose $u,v\in V$ are the sides of the rhombus and $\norm{u} = \norm{v}$. 
We can directly apply our result from part (b) to state $\langle u+v,u-v\rangle=0$, implying that the diagonals of the rhombus are orthogonal, and thus, perpendicular to each other.

\clearpage

\section{No eigenvalues greater than 1}
\subsection*{Problem statement}
Suppose $T\in \mathcal{L}(V)$ is such that $\norm{Tv}\leq\norm{v}$ for every $v\in V$. 
Prove that $T-\sqrt{2}I$ is invertible.

\subsection*{Solution}
Theorem 5.6 (`Equivalent conditions to be an eigenvalue') implies that ``$T-\sqrt{2}I$ is invertible'' means that $\sqrt{2}$ is not an eigenvalue of $T$. 
Let's first note that $1<\sqrt{2}$. 

Suppose $v$ is an eigenvector of $T$. 
The condition that $\norm{Tv}\leq\norm{v}$ implies that the eigenvalue corresponding to $v$, let's say $\lambda$, we have
\[\abs{\lambda}\norm{v}\leq\norm{v},\]
and thus
\[\lambda\leq 1.\]
Hence, $\sqrt{2}$ cannot be an eigenvalue and $T-\sqrt{2}I$ is invertible.

\clearpage

\section{Condition equivalent to orthogonality}
\subsection*{Problem statement}
Suppose $u,v\in V$. 
Prove that $\langle u,v\rangle=0$ if and only if
\[\norm{u}\leq\norm{u+av}\]
for all $a\in\mathbf{F}$.

\subsection*{Solution}
\subsubsection*{First Direction}
Suppose $\langle u,v\rangle=0$. 
It follows that for all $a\in\mathbf{F}$, we have
\[\langle u,av\rangle=0.\]
Via the Pythagorean Theorem (Theorem 6.13), we can write
\[\norm{u}^2+\norm{av}^2=\norm{u+av}^2.\]
Given the positivity condition of inner products (Definition 6.3), it follows that
\[\norm{u}^2\leq\norm{u+av}^2.\]

\subsubsection*{Second Direction}
Suppose $\norm{u}\leq\norm{u+av}$ for all $a\in\mathbf{F}$. 
Via Theorem 6.14 (`An orthogonal decomposition'), we can decompose $u$ into $cv$ and $w$ where $c=\frac{\langle u,v\rangle}{\norm{v}^2}$, $w=u-cv$, and $\langle cv,w\rangle=0$. 
Invoking the Pythagorean Theorem (Theorem 6.13), we can write
\[\norm{u}^2=\norm{cv+w}^2=\norm{cv}^2+\norm{w}^2=\norm{cv}^2+\norm{u-cv}^2,\]
and via the positivity condition of inner products (Definition 6.3), it follows that
\[\norm{u}^2\geq\norm{u-cv}^2\]
and by taking the square root of both sides
\[\norm{u}\geq\norm{u-cv}.\]
In order to avoid a contradiction with our hypothesis, the expression above must be an equality. 
The expression above can only be an equality if $c=0$ or $v=0$. 
Both cases imply that $\langle u,v\rangle=0$.

\clearpage

\section{$\norm{au+bv}=\norm{bu+av}$ iff $\norm{u}=\norm{v}$}
\subsection*{Problem statement}
Suppose $u,v\in V$. 
Prove that $\norm{au+bv}=\norm{bu+av}$ for all $a,b\in \mathbf{R}$ if and only if $\norm{u}=\norm{v}$.

\subsection*{Solution}
As an organizational note, it's easier to work with squared norms, and the results directly translate to norms. 

\subsubsection*{First Direction}
Suppose $\norm{au+bv}^2=\norm{bu+av}^2$ for all $a,b\in \mathbf{R}$. 
Expanding out $\norm{au+bv}^2=\norm{bu+av}^2$, we have
\[\langle au,au\rangle+\langle bv,au\rangle + \langle au,bv\rangle + \langle bv,bv\rangle=\langle bu,bu\rangle+\langle av,bu\rangle + \langle bu,av\rangle + \langle av,av\rangle.\]
Since $a,b\in\mathbf{R}$, it follows that
\[\langle bv,au\rangle=b\bar{a}\langle v,u\rangle=ab\langle v,u\rangle,\]
and a similar result follows for $\langle au,bv\rangle,\langle av,bu\rangle,\langle bu,av\rangle$.
Hence, we can write
\[a^2\norm{v}^2+ab\langle v,u\rangle + ab\langle u,v\rangle + b^2\norm{u}^2=b^2\norm{v}^2+ab\langle v,u\rangle + ab\langle u,v\rangle + a^2\norm{u}^2,\]
and by subtracting out the similar terms, we have
\[a^2\norm{v}^2 + b^2\norm{u}^2=b^2\norm{v}^2 + a^2\norm{u}^2,\]
By setting $b=0$, it follows that
\[a^2\norm{v}^2=a^2\norm{u}^2,\]
which implies $\norm{u}^2=\norm{v}^2$ and $\norm{u}=\norm{v}$.

\subsubsection*{First Direction}
Suppose $\norm{u}^2=\norm{v}^2$. 
It follows that for all $a,b\in \mathbf{R}$, 
\[a^2\norm{v}^2=a^2\norm{u}^2 \;\;\;\text{and}\;\;\; b^2\norm{v}^2=b^2\norm{u}^2.\]
By adding the two expressions together, we have
\[a^2\norm{v}^2 + b^2\norm{u}^2=b^2\norm{v}^2 + a^2\norm{u}^2.\]
To construct $\norm{au+bv}^2=\norm{bu+av}^2$, one only needs to reverse our work in the \textbf{First Direction}.

\clearpage

\section{Properties of $u=v$}
\subsection*{Problem statement}
Suppose $u,v\in V$ and $\norm{u}=\norm{v}=1$ and $\langle u,v\rangle=1$. 
Prove that $u=v$.

\subsection*{Solution}
Via the Cauchy-Schwarz Inequality (Theorem 6.15), we know that
\[\abs{\langle u,v\rangle}\leq\norm{u}\norm{v},\]
where the inequality is an equality if and only if one of $u,v$ is a scalar multiple of the other. 
In our case, we have
\[1=\langle u,v\rangle = \norm{u}\norm{v}=1\cdot 1=1,\]
hence, we have an equality and $u=cv$ for some $c\in\mathbf{F}$. 
To compute the exact $c$, Theorem 6.14 (`An orthogonal decomposition') tells us that 
\[c=\frac{\langle u,v\rangle}{\norm{v}^2}=\frac{1}{1^2}=1.\]
Therefore, $u=v$.

\clearpage

\section{When $\norm{u}\leq 1$ and $\norm{v}\leq 1$}
\subsection*{Problem statement}
Suppose $u,v\in V$ and $\norm{u}\leq 1$ and $\norm{v}\leq 1$. 
Prove that
\[\sqrt{1-\norm{u}^2}\sqrt{1-\norm{v}^2}\leq 1-\abs{\langle u,v \rangle}.\]

\subsection*{Solution}
Via the Cauchy-Schwarz Inequality (Theorem 6.15) and the problem statement, we have
\[\abs{\langle u,v\rangle}\leq \norm{u}\norm{v}\leq 1.\]
By subtracting by $1$ and multiplying by $-1$, we have
\[0\leq 1-\norm{u}\norm{v}\leq 1- \abs{\langle u,v\rangle}.\]
Thus, if we can show that $\sqrt{1-\norm{u}^2}\sqrt{1-\norm{v}^2}\leq1-\norm{u}\norm{v}$, then our desired result immediately follows.

By squaring the expression $\sqrt{1-\norm{u}^2}\sqrt{1-\norm{v}^2}\leq 1-\norm{u}\norm{v}$ on both sides (which is allowed since both terms are real and greater than or equal to $0$), we have
\[(1-\norm{u}^2)(1-\norm{v}^2)\leq (1-\norm{u}\norm{v})^2\]
which can be expanded to
\[1-\norm{u}^2-\norm{v}^2+\norm{u}^2\norm{v}^2\leq 1-2\norm{u}\norm{v}+\norm{u}^2\norm{v}^2\]
and then reduced to
\[-\norm{u}^2-\norm{v}^2 \leq -2\norm{u}\norm{v}.\]
By rearranging terms, we can write
\[0\leq \norm{u}^2 - 2\norm{u}\norm{v} + \norm{v}^2,\]
and it follows that
\[0\leq (\norm{u} - \norm{v})^2\]
which, given the positivity property of inner products (Definition 6.3), is clearly true. 
Therefore, we can state that 
\[\sqrt{1-\norm{u}^2}\sqrt{1-\norm{v}^2}\leq 1-\norm{u}\norm{v}\]
and our desired result immediately follows.

\clearpage

\section{Searching for vectors in $\mathbf{R}^2$}
\subsection*{Problem statement}
Find vectors $u,v\in \mathbf{R}^2$ such that $u$ is a scalar multiple of $(1,3)$, $v$ is orthogonal to $(1,3)$, and $(1,2)=u+v$.

\subsection*{Solution}
First, let's find a vector orthogonal to $(1,3)$. 
By inspection, the vector $(-3,1)$ will work. 
Now we need to find $a,b\in\mathbf{R}$ such that $u=a(1,3)$, $v=b(-3,1)$, and $(1,2)=u+v$. 
Using the constraint $(1,2)=u+v$, we get the following equations
\begin{align*} 
1=a-3b &\Rightarrow 3a=9b+3, \\ 
2=3a+b &\Rightarrow 3a=2-b,
\end{align*}
and through combining them, we have
\[9b+3=2-b\]
and $b=-\frac{1}{10}$. 
Plugging in $b$, we have $a=\frac{7}{10}$. 
Hence, our vectors are
\[u=(\frac{7}{10},\frac{21}{10}) \;\;\;\text{and}\;\;\; v=(\frac{3}{10},-\frac{1}{10}).\]

\end{document}